From the function-sheaf dictionary to quasicharacters of $p$-adic tori

### 2015-08-03 dictionary.tex
Re-added the commented-out paragraph in the proof of the classical dictionary.

\begin{proposition}\label{prop:connected}
 If $G$ is a connected, commutative algebraic group over $\Fq$ then:
 \begin{enumerate}
 \labitem{(1)}{c1} $\TrFrob{G} : \CSiso{G} \to G(\Fq)^*$ is an isomorphism of groups;
 \labitem{(2)}{c2} every character sheaf on $G$ is isomorphic to one defined by a discrete isogeny;
 \labitem{(3)}{c3} $\Aut(\cs{L}) = 1$, for all character sheaves $\cs{L}$ on $G$.
 \end{enumerate}
 \end{proposition}
\begin{proof}
By Lemma~\ref{lem:bounded_connected}, we know that every character sheaf $\cs{L}$ on $\bG$ descends to $G$;
let $\cs{E}$ be an object of $\bCS(G)$ for which $B_G(\cs{E}) \iso \cs{L}$.
 Since the functor $B_G : \bCS(G) \to \CS(G)$ is full and faithful, $\Aut(\cs{L}) = \Aut(\cs{E})$.
From here, Deligne's function-sheaf dictionary for connected commutative algebraic groups over finite fields,
as in \cite{deligne:SGA4.5}*{Sommes trig.} or \cite{laumon:87a}*{1.1.3}, gives us all we need for points \ref{c1} and \ref{c2}.
\iffalse
As in the proof of Proposition~\ref{prop:finite}, use \'etale descent to see that pullback by the Lang isogeny $\Lang : G\to G$
defines an equivalence of categories between local systems on $G$ and $G(\Fq)$-equivariant local systems on $G$. 
Under this equivalence, local systems $\cs{E}$ on $G$ arising from objects in $\bCS(G)$ are matched with $G(\Fq)$-equivariant
constant local systems of rank-one on $G$, and therefore with one-dimensional representations of $G(\Fq)$. 
In the same way, pullback along the isogeny $\Lang\times\Lang : G\times G\to G\times G$ matches the extra structure
$\mu_\cs{E} : m^*\cs{E} \to \cs{E}\boxtimes\cs{E}$ with an isomorphism $m^*V \to V\boxtimes V$ of one-dimensional
representations of $G(\Fq)\times G(\Fq)$, which is exactly an isomorphism $V \to V\otimes V$ of one-dimensional representations,
which is exactly the choice of an isomorphism $V\iso \EE$.
We see that $\bCS(G)$ is equivalent to the category of characters of $G(\Fq)$.
Let $\bg$ be a geometric point above $g \in G(\Fq)$.  If $\cs{E}$ matches $\psi : G(\Fq)\to \EEx$ under this equivalence,
a simple calculation on stalks reveals that the action of Frobenius on $\cs{E}_{\bg}$ is multiplication by $\psi(g)^{-1}$.
In other words, for every $\cs{E}$ in $\bCS(G)$, the trace of $\Lang^*\cs{E}$ is $\trFrob{\cs{E}}^{-1}$ as a
representation of $G(\Fq)$, proving parts \ref{c1} and \ref{c2}.
\fi
For part \ref{c3}, suppose $\Lang^*\cs{E} = V$ with isomorphism $V \to V\otimes V$.
Observe that the equivalence above establishes a bijection between $\Aut(\cs{E})$ and the group of automorphisms of $\rho : V\to V$ for which 
\[
\begin{tikzcd}
\arrow{d}{} V \arrow{r}{\rho} & V\arrow{d}{}\\
V\otimes V \arrow{r}{\rho\otimes \rho} & V\otimes V
\end{tikzcd}
\]
commutes. 
Since the only such isomorphism $\rho$ is $\id_V$, it follows that $\Aut(\cs{E}) = 1$, completing the proof.
\end{proof}

>>>>

\begin{proposition}\label{prop:connected}
 If $G$ is a connected, commutative algebraic group over $\Fq$ then:
 \begin{enumerate}
 \labitem{(1)}{c1} $\TrFrob{G} : \CSiso{G} \to G(\Fq)^*$ is an isomorphism of groups;
 \labitem{(2)}{c2} every character sheaf on $G$ is isomorphic to one defined by a discrete isogeny;
 \labitem{(3)}{c3} $\Aut(\cs{L}) = 1$, for all character sheaves $\cs{L}$ on $G$.
 \end{enumerate}
 \end{proposition}
\begin{proof}
By Lemma~\ref{lem:bounded_connected}, we know that every character sheaf $\cs{L}$ on $\bG$ descends to $G$;
let $\cs{E}$ be an object of $\bCS(G)$ for which $B_G(\cs{E}) \iso \cs{L}$.
 Since the functor $B_G : \bCS(G) \to \CS(G)$ is full and faithful, $\Aut(\cs{L}) = \Aut(\cs{E})$.
From here, Deligne's function-sheaf dictionary for connected commutative algebraic groups over finite fields,
as in \cite{deligne:SGA4.5}*{Sommes trig.} or \cite{laumon:87a}*{1.1.3}, gives us all we need for points \ref{c1} and \ref{c2}, as we briefly recall.

%\iffalse
As in the proof of Proposition~\ref{prop:finite}, use \'etale descent to see that pullback by the Lang isogeny $\Lang : G\to G$
defines an equivalence of categories between local systems on $G$ and $G(\Fq)$-equivariant local systems on $G$. 
Under this equivalence, local systems $\cs{E}$ on $G$ arising from objects in $\bCS(G)$ are matched with $G(\Fq)$-equivariant
constant local systems of rank-one on $G$, and therefore with one-dimensional representations of $G(\Fq)$. 
In the same way, pullback along the isogeny $\Lang\times\Lang : G\times G\to G\times G$ matches the extra structure
$\mu_\cs{E} : m^*\cs{E} \to \cs{E}\boxtimes\cs{E}$ with an isomorphism $m^*V \to V\boxtimes V$ of one-dimensional
representations of $G(\Fq)\times G(\Fq)$, which is exactly an isomorphism $V \to V\otimes V$ of one-dimensional representations,
which is exactly the choice of an isomorphism $V\iso \EE$.
We see that $\bCS(G)$ is equivalent to the category of characters of $G(\Fq)$.
Let $\bg$ be a geometric point above $g \in G(\Fq)$.  If $\cs{E}$ matches $\psi : G(\Fq)\to \EEx$ under this equivalence,
a simple calculation on stalks reveals that the action of Frobenius on $\cs{E}_{\bg}$ is multiplication by $\psi(g)^{-1}$.
In other words, for every $\cs{E}$ in $\bCS(G)$, the trace of $\Lang^*\cs{E}$ is $\trFrob{\cs{E}}^{-1}$ as a
representation of $G(\Fq)$, proving parts \ref{c1} and \ref{c2}.
%\fi

For part \ref{c3}, suppose $\Lang^*\cs{E} = V$ with isomorphism $V \to V\otimes V$.
Observe that the equivalence above establishes a bijection between $\Aut(\cs{E})$ and the group of automorphisms of $\rho : V\to V$ for which 
\[
\begin{tikzcd}
\arrow{d}{} V \arrow{r}{\rho} & V\arrow{d}{}\\
V\otimes V \arrow{r}{\rho\otimes \rho} & V\otimes V
\end{tikzcd}
\]
commutes. 
Since the only such isomorphism $\rho$ is $\id_V$, it follows that $\Aut(\cs{E}) = 1$, completing the proof.
\end{proof}
